\newgeometry{left=35mm,right=25mm,top=35mm,bottom=35mm}
\begin{center}
{\Huge \bf Acknowledgements}
\end{center}
\noindent
  % Pete -->
  I am enormously grateful to my supervisor, Pete Philipson, for his unwavering support, empathy, and patience over my studies. His guidance helped me develop as a researcher, and I'll miss our weekly meetings filled with debugging, rewriting and having a good natter. Additionally, Pete was immensely supportive through a period of uncertainty in my life, offering great advice and remaining a stalwart source of encouragement, for which I'm especially thankful.\\ Many thanks to Dr.\ Joe Matthews and Dr.\ Lisa McFetridge for examining this thesis.

  % Friends / AGE / Mum n Dad -->
  I have been lucky enough to make many friends within the department, whether this was through sharing an office or eating lunch together, they made the department a great place to work: Ryan Doran, Thomas Flynn, Nicola Hewett, Sam Hartharn-Evans, Nick Keepfer, Kieran Peel and Tom Billam. I'd also like to thank the fantastic computing officer Michael Beaty, who the School is lucky to have, for putting up with many inane queries from me.\newline
  Outside of the department I'd like to thank my friends Eve and Aaron, Charlotte, as well as Anna, Sophie, India, and Laura. Last, but obviously not least, I'm forever grateful to my Mum and Dad for their unwavering support and encouragement throughout my journey in education, for looking after Beans, and much more.

  % Funding -->
  I would also like to thank the Engineering and Physical Sciences Research Council for funding my PhD studies (grant reference EP/V520184/1).
    
\thispagestyle{empty}
\restoregeometry